Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/evaluation.tex
spell

\vspace{.5em}\noindent\textbf{TPC-C} To better understand the impact
of HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark. Surprisingly, four of five
transactions can be executed with HATs, while the fifth may require
unavailablity.

TPC-C consists of five transactions, capturing the operation of a
wholesale warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely under HATs. Clients may read
stale data, but this does not violate TPC-C requirements and clients
will read their writes if they are sticky-available. Another
transaction type, \textit{Payment}, updates running balances for
warehouses, districts, and customer records as well as providing an
audit trail. The transaction is increment- and append-only, so all
balance increase operations commute, and TA allows the
maintainence of any foreign-key constraints (e.g.,
\texttt{UPDATE/DELETE CASCADE}).

@@

New-Order presents two challenges: ID assignment and stock
maintainence. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with much more stock
than in a non-HAT-compliant implementation.

@@

Throughout execution, TPC-C also requires the maintainence of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via transactional atomicity, can be easily maintained. Consistency
Conditions 4 through 12 (3.3.2.4-12) can similarly be satisfied by
applying updates atomically across tables. Consistency Conditions 2
and 3 (3.3.2.2-3) concern order ID assignment and are more
problematic, as we have discussed.  Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests, such as one that requires simultaneously modifying a product
description and its image, are also passable using HATs.

>>>>

\vspace{.5em}\noindent\textbf{TPC-C} To better understand the impact
of HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark. Surprisingly, four of five
transactions can be executed with HATs, while the fifth may require
unavailability.

TPC-C consists of five transactions, capturing the operation of a
wholesale warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely under HATs. Clients may read
stale data, but this does not violate TPC-C requirements and clients
will read their writes if they are sticky-available. Another
transaction type, \textit{Payment}, updates running balances for
warehouses, districts, and customer records as well as providing an
audit trail. The transaction is increment- and append-only, so all
balance increase operations commute, and TA allows the
maintenance of any foreign-key constraints (e.g.,
\texttt{UPDATE/DELETE CASCADE}).

@@

New-Order presents two challenges: ID assignment and stock
maintenance. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with much more stock
than in a non-HAT-compliant implementation.

@@

Throughout execution, TPC-C also requires the maintenance of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via transactional atomicity, can be easily maintained. Consistency
Conditions 4 through 12 (3.3.2.4-12) can similarly be satisfied by
applying updates atomically across tables. Consistency Conditions 2
and 3 (3.3.2.2-3) concern order ID assignment and are more
problematic, as we have discussed.  Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests, such as one that requires simultaneously modifying a product
description and its image, are also passable using HATs.

### 2013-04-01 writeup/model.tex
spell

Traditionally, a system provides {\em availability} if every user that
can contact a correct server eventually receives a response from that
server, even in the presence of arbitrary, indefinitely long network
partitions between servers~\cite{gilbert-cap}.\footnote{Under this
  definition, a system that requires a majority of servers to be
  online is not availabile.} As in a standard distributed database,
designated servers might handle operations on different data
items;\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{dynamo,
    pnuts, bigtable, spanner, hstore}, we consider partially
  replicated systems.} a server that can handle an operation for a
given data item is called a \textit{replica} for that item. While this
standard model is useful, there are several other cases we should
consider.

>>>>

Traditionally, a system provides {\em availability} if every user that
can contact a correct server eventually receives a response from that
server, even in the presence of arbitrary, indefinitely long network
partitions between servers~\cite{gilbert-cap}.\footnote{Under this
  definition, a system that requires a majority of servers to be
  online is not available.} As in a standard distributed database,
designated servers might handle operations on different data
items;\footnote{There is a further distinction between a \textit{fully
    replicated} system, in which all servers are replicas for all data
  items, and a \textit{partially replicated} system, in which at least
  one server acts as a replica for only a subset of all data
  items. For generality, and, for applicability to many modern
  ``sharded'' or ``partitioned'' data storage systems~\cite{dynamo,
    pnuts, bigtable, spanner, hstore}, we consider partially
  replicated systems.} a server that can handle an operation for a
given data item is called a \textit{replica} for that item. While this
standard model is useful, there are several other cases we should
consider.

### 2013-04-01 writeup/motivation.tex
spell

Even with fault-free networks, distributed systems face the challenge
of communication latency, Deutsch's second ``Fallacy.'' In this
section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds---in a geo-replicated context.
Fundamentally, the speed at which two servers can communicate is
(according to modern physics) bounded by the speed of light. In the
best case, two servers on opposite sides of the Earth's
surface---communicating via a hypothetical link traversing the
planet's diamater---require a minimum 85.1 ms RTT (133.7 ms if sent at
ground level). As services are replicated to multiple, geographically
distinct sites, the cost of communication increases to this lower
bound.

>>>>

Even with fault-free networks, distributed systems face the challenge
of communication latency, Deutsch's second ``Fallacy.'' In this
section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds---in a geo-replicated context.
Fundamentally, the speed at which two servers can communicate is
(according to modern physics) bounded by the speed of light. In the
best case, two servers on opposite sides of the Earth's
surface---communicating via a hypothetical link traversing the
planet's diameter---require a minimum 85.1 ms RTT (133.7 ms if sent at
ground level). As services are replicated to multiple, geographically
distinct sites, the cost of communication increases to this lower
bound.
